On the assessment of the ability of measurements, nowcasts, and forecasts to track changes

### 2024-08-01 paper.tex
Remove comments
Submitted version
Adjust paper formatting according to journal instructions

%\documentclass[referee,sn-mathphys-ay,pdflatex]{sn-jnl}
%\documentclass[sn-mathphys-ay,pdflatex]{sn-jnl}
\documentclass[pdflatex]{sn-jnl}
% Language setting
% Replace `englis' with e.g. `spanish' to change the document language
\usepackage[english]{babel}
\usepackage[utf8]{inputenc} % fuer Umlaute
\usepackage{fontenc}

% Set page size and margins
% Replace `letterpaper' with`a4paper' for UK/EU standard size
%\usepackage[a4paper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

%%%% ---- Useful packages

@@

%%% --- Tabulars
\usepackage{booktabs}
\usepackage{multirow} % cells over several rows
\usepackage{tabularx}
%\usepackage{tabularray}
\usepackage{floatrow}
\DeclareFloatFont{tiny}{\tiny}% "scriptsize" is defined by floatrow, "tiny" not
\floatsetup[table]{font=scriptsize} %https://tex.stackexchange.com/a/27100

%%% --- Mathematics
\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools, mathabx}
\usepackage{bm} % bold mathematics
\usepackage{bbm} % for blackboard 1
\usepackage[linesnumbered,lined,algo2e,figure,boxed]{algorithm2e}
\usepackage[short]{optidef} % for aligning linear programs
%\usepackage{nccmath} % for fleqn environment

%%% --- Other
%\usepackage[colorlinks=true, allcolors=blue]{hyperref}
%\usepackage[hidelinks]{hyperref}
\usepackage[style=authoryear,backend=biber,natbib=true]{biblatex}
%\usepackage{natbib}
%\bibliographystyle{
%\addbibresource{library.bib}
\addbibresource{library-jonas.bib}
\usepackage{csquotes}
\usepackage[shortcuts]{extdash} % For Hyphens in English
%\usepackage{acro}
% \DeclareAcronym{pdf}{short=PDF, long={probability distribution function}}

%%% --- Commenting, Todonotes
%\usepackage[colorinlistoftodos,prependcaption]{todonotes}
%%\usepackage{comment, soul}
%%\newcommand{\ar}{$\rightarrow$}
%\presetkeys{todonotes}{backgroundcolor=yellow!30, bordercolor=yellow!50, linecolor=yellow!50, figwidth=\textwidth}{}
%\newcommand{\hl}[1]{\textcolor{Aquamarine}{#1}}

%\usepackage{xargs,ifthen} % Use optional arguements in commands
%\newcommandx{\unsure}[2][1=]{\todo[linecolor=blue,backgroundcolor=blue!25,bordercolor=blue,#1]{#2}}
%\newcommandx{\discuss}[2][1=]{\todo[linecolor=orange,backgroundcolor=orange!25,bordercolor=orange,#1]{#2}}
%\newcommandx{\missing}[2][1=]{\todo[linecolor=red,backgroundcolor=red!25,bordercolor=red,#1]{#2}}
%\newcommandx{\change}[2][1=]{\todo[linecolor=red,backgroundcolor=red!25,bordercolor=red,#1]{#2}}
%\newcommandx{\info}[2][1=]{\todo[linecolor=OliveGreen,backgroundcolor=OliveGreen!25,bordercolor=OliveGreen,#1]{#2}}
%\newcommandx{\improvement}[2][1=]{\todo[linecolor=Plum,backgroundcolor=Plum!25,bordercolor=Plum,#1]{#2}}
%\newcommandx{\done}[2][1=]{\todo[linecolor=gray,backgroundcolor=gray!25,bordercolor=gray,#1]{#2}}
%\newcommandx{\thiswillnotshow}[2][1=]{\todo[disable,#1]{#2}}
%\usepackage{marginnote}
%\let\marginpar\marginnote

@@

\newcommand{\acc}{\mu}
\newcommand{\accp}{\acc^+}
\newcommand{\accm}{\acc^-}
%\newcommand{\acceps}[1][]{\ifthenelse { \equal {#1} {} }{\acc_{\eps}}{\acc_{#1}}} %https://stackoverflow.com/a/7314951
\newcommand{\acceps}[1][e]{\acc_{#1}} %https://stackoverflow.com/a/7314951
\newcommand{\accpeps}[1][e]{\acceps[#1]^+}
\newcommand{\accmeps}[1][e]{\acceps[#1]^-}
\newcommand{\accl}[1][l]{\mu^{#1}}
\newcommand{\accpl}[1][l]{\acc^{+,#1}}
\newcommand{\accml}[1][l]{\acc^{-,#1}}

@@

% Ungefähre Wortanzahl des Hauptteils (von Introduction bis Conclusion): oben links auf 'Menu' und dann 'Word Count'
% Ungefähre Aufteilung auf Abschnitte (nicht direkt in Overleaf sichtbar):
% Introduction: 500
% Assessment of ATC: 2225
% Application: 3045
% Discussion and Conclusion: 587

% Abstract: 99

%%% --- ToDo List
%\usepackage{enumitem}
%\newlist{todolist}{itemize}{2}
%\setlist[todolist]{label=$\square$}
%\usepackage{pifont}
%\newcommand{\cmark}{\ding{51}}%
%\newcommand{\xmark}{\ding{55}}%
%\newcommand{\done}{\rlap{$\square$}{\raisebox{2pt}{\large\hspace{1pt}\cmark}}%
%\hspace{-2.5pt}}
%\newcommand{\wontfix}{\rlap{$\square$}{\large\hspace{1pt}\xmark}}

@@

\title{On the assessment of the ability of measurements, nowcasts, and forecasts to track changes}
%\author{Oliver Grothe, Bolin Liu, Jonas Rieger}
\author*[1]{Jonas Rieger}\email{jonas.rieger@kit.edu}
\author[1]{Bolin Liu}
\author[2]{Bernd Saugel}
\author[1]{Oliver Grothe}

\affil[1]{\orgdiv{Institute for Operations Research}, \orgname{Karlsruhe Institute of Technology (KIT)}, \orgaddress{\street{Kaiserstr. 12}, \city{Karlsruhe}, \postcode{76131}, \country{Germany}}}
\affil[2]{\orgdiv{Department of Anesthesiology, Center of Anesthesiology and Intensive Care Medicine}, \orgname{University Medical Center Hamburg-Eppendorf}, \orgaddress{\street{Martinistrasse 52}, \city{Hamburg}, \postcode{20246}, \country{Germany}}}
%TC:ignore
\abstract{ 
Measurements, nowcasts, or forecasts ideally should correctly reflect changes in the values of interest.
In this article, we focus on how to assess the ability of measurements, nowcasts, or forecasts to correctly predict the direction of changes in values – which we refer to as the ability to track changes. 
We present visual techniques and quantitative measures to assess the ability to track changes and provide extensions for noisy data and estimation uncertainty using bootstrap confidence intervals and exclusion areas.
We exemplarily illustrate the proposed methods to assess the ability to track changes for nowcasting during the COVID-19 pandemic, patient admissions to an emergency department, and non-invasive blood pressure measurements.
}
%TC:endignore
\keywords{}

\maketitle
% \listoftodos

@@

The estimators accept various shapes of the exclusion area (see Figure~\ref{fig:aatc_4q}).
A rectangular exclusion area, $E = \{(y, x) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x) \land (-\varepsilon_y \leq y \leq \varepsilon_y) \}$ for $\varepsilon_x, \varepsilon_y > 0$, leaves out points that are small in both components.
%Points are included if the true value or prediction is unlikely to be noise-driven.
%In the example graph in Figure~\ref{fig:aatc_basic_4q_excl_box}, only point 1 is excluded.
An exclusion area along one axis, for example, $E = \{(y, x) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x)\}$ for $\varepsilon_x > 0$, removes points in which one of the components could change sign by a small amount of noise.
%At least for measurement, nowcast, or forecast, small amounts of noise are inevitable.
A cross-shaped exclusion area, $E = \{(y, x) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x) \lor (-\varepsilon_y \leq y \leq \varepsilon_y) \}$ for $\varepsilon_x, \varepsilon_y > 0$, along both axes accounts for the sign reversal in both components.
%A point is only included if no component is likely to be driven by noise.
%For the latter two methods, points 1 and 7 in Figure~\ref{fig:aatc_basic_4q_excl_axis} or 1, 6, and 7 are excluded in Figure~\ref{fig:aatc_basic_4q_excl_cross}, respectively.

@@

\begin{figure}
    \centering
%    \includegraphics{}
    \begin{subfigure}[t]{.48\textwidth}
    \includegraphics{plots/covid_nowcast/40_cond_prob_lag_7}
    \caption{Conditional ATC plot.}\label{fig:app-covid-cond-prob-7}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{.48\textwidth}
    \includegraphics{plots/covid_nowcast/40_acc_eps_lag_7}
    \caption{ATC ratio over exclusion area size in $\diffx$.}\label{fig:app-covid-atc-ratio-7}
    \end{subfigure}
    \caption[Conditional ATC plot and ATC ratio over exclusion area for the COVID-19 nowcasts.]{Conditional \ac{atc} plot and \ac{atc} ratio over exclusion area for the nowcasts of the seven-day hospitalization rate ILM, RKI, RIVM, and ENS-MED for the horizon seven days in COVID-19 nowcasting.}
    \label{fig:app-covid-cond-prob-atc-ratio-7}
\end{figure}

@@

%TC:ignore
\section*{Declarations}
\subsection*{Availability of data and materials}

The datasets analyzed during the study are available in the corresponding databases or repositories \parencite{ChairOfEconometricsAndStatisticsAtKarlsruheInstituteOfTechnology2024, Rostami-Tabar2023, Moody2017}.
All code is available in the repository [GITHUB LINK WILL BE PROVIDED AFTER REVIEW].

@@

\printbibliography
%\bibliography{library,library-jonas}

@@

\begin{figure}
    \centering
%    \includegraphics{}
    \begin{subfigure}[t]{.48\textwidth}
    \includegraphics{plots/covid_nowcast/40_cond_prob_lag_1}
    \caption{Conditional \ac{atc} plot for horizon one.}\label{fig:app-covid-cond-prob-1}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{.48\textwidth}
    \includegraphics{plots/covid_nowcast/40_cond_prob_lag_14}
    \caption{Conditional \ac{atc} plot for horizon 14.}\label{fig:app-covid-cond-prob-14}
    \end{subfigure}
    \begin{subfigure}[t]{.48\textwidth}
    \includegraphics{plots/covid_nowcast/40_acc_eps_lag_1}
    \caption{\Ac{atc} ratio over exclusion area size in $\diffx$ for horizon one.}\label{fig:app-covid-atc-ratio-1}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{.48\textwidth}
    \includegraphics{plots/covid_nowcast/40_acc_eps_lag_14}
    \caption{\ac{atc} ratio over exclusion area size in $\diffx$ for horizon 14.}\label{fig:app-covid-atc-ratio-14}
    \end{subfigure}
    \caption[Conditional ATC plot and ATC ratio over exclusion area in COVID-19 nowcasting.]{Conditional \ac{atc} plot and \ac{atc} ratio over exclusion area for the COVID-19 nowcast models ILM, RKI, RIVM, and ENS-MED for the horizon seven days. Note the different axis scalings of the ATC ratios for the two horizons. While the ENS-MED model has a rather low \ac{atc} ratio for small exclusion areas, it is increasing fast for larger exclusion areas. The RKI model is among the worst, except for the small negative predicted changes. }
    \label{fig:app-covid-cond-prob-atc-ratio-1-14}
\end{figure}

\begin{figure}
    \centering
    \begin{subfigure}[t]{.48\textwidth}
        \includegraphics{plots/covid_nowcast/60_reliability_diagram_lag_7}
        \caption{Reliability diagram for horizon seven days.} \label{fig:app-covid-reliability-7}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{.48\textwidth}
        \includegraphics{plots/covid_nowcast/60_reliability_diagram_lag_14}
        \caption{Reliability diagram for horizon 14 days.} \label{fig:app-covid-reliability-14}
    \end{subfigure}
    \begin{subfigure}{\textwidth}
        \includegraphics{plots/covid_nowcast/70_prob_hist}
        \caption{Count histogram of the predicted probabilities for the horizon one, seven, and 14 days.} \label{fig:app-covid-prob-hist}
    \end{subfigure}
    \caption[Reliability diagrams for COVID-19 nowcasting models and the horizon seven and 14 days.]{The reliability diagram for the COVID-19 nowcasting models ILM, RIVM, RKI, and ENS-MED for the horizon seven and 14 days.
Additionally, the count of predicted probabilities for the horizons is shown.
    The reliability diagram bins are chosen according to the empirical quantiles of the predicted probabilities.
As the models issue small or large probabilities of increase for the higher horizons, little information on the accuracy of moderate probability predictions is available.}
    \label{fig:app-covid-rel}
\end{figure}
%TC:endignore

>>>>

\documentclass[pdflatex]{sn-jnl}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc} % fuer Umlaute
\usepackage{fontenc}

@@

%%% --- Tabulars
\usepackage{booktabs}
\usepackage{multirow} % cells over several rows
\usepackage{tabularx}
\usepackage{floatrow}
\DeclareFloatFont{tiny}{\tiny}% "scriptsize" is defined by floatrow, "tiny" not
\floatsetup[table]{font=scriptsize} %https://tex.stackexchange.com/a/27100

%%% --- Mathematics
\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools, mathabx}
\usepackage{bm} % bold mathematics
\usepackage{bbm} % for blackboard 1
\usepackage[linesnumbered,lined,algo2e,figure,boxed]{algorithm2e}
\usepackage[short]{optidef} % for aligning linear programs

%%% --- Other
\usepackage[style=numeric,backend=biber,natbib=true]{biblatex}
\addbibresource{paper_biber.bib}
\usepackage{csquotes}
\usepackage[shortcuts]{extdash} % For Hyphens in English

@@

\newcommand{\acc}{\mu}
\newcommand{\accp}{\acc^+}
\newcommand{\accm}{\acc^-}
\newcommand{\acceps}[1][e]{\acc_{#1}} %https://stackoverflow.com/a/7314951
\newcommand{\accpeps}[1][e]{\acceps[#1]^+}
\newcommand{\accmeps}[1][e]{\acceps[#1]^-}
\newcommand{\accl}[1][l]{\mu^{#1}}
\newcommand{\accpl}[1][l]{\acc^{+,#1}}
\newcommand{\accml}[1][l]{\acc^{-,#1}}

@@

@@

\title{On the assessment of the ability of measurements, nowcasts, and forecasts to track changes}
\author*[1]{Jonas Rieger}\email{jonas.rieger@kit.edu}
\author[1]{Bolin Liu}
\author[2]{Bernd Saugel}
\author[1]{Oliver Grothe}

\affil[1]{\orgdiv{Institute for Operations Research}, \orgname{Karlsruhe Institute of Technology (KIT)}, \orgaddress{\street{Kaiserstr. 12}, \city{Karlsruhe}, \postcode{76131}, \country{Germany}}}
\affil[2]{\orgdiv{Department of Anesthesiology, Center of Anesthesiology and Intensive Care Medicine}, \orgname{University Medical Center Hamburg-Eppendorf}, \orgaddress{\street{Martinistrasse 52}, \city{Hamburg}, \postcode{20246}, \country{Germany}}}
\abstract{
Measurements, nowcasts, or forecasts ideally should correctly reflect changes in the values of interest.
In this article, we focus on how to assess the ability of measurements, nowcasts, or forecasts to correctly predict the direction of changes in values – which we refer to as the ability to track changes. 
We present visual techniques and quantitative measures to assess the ability to track changes and provide extensions for noisy data and estimation uncertainty using bootstrap confidence intervals and exclusion areas.
We exemplarily illustrate the proposed methods to assess the ability to track changes for nowcasting during the COVID-19 pandemic, patient admissions to an emergency department, and non-invasive blood pressure measurements.
}
\keywords{}

\maketitle

@@

The estimators accept various shapes of the exclusion area (see Figure~\ref{fig:aatc_4q}).
A rectangular exclusion area, $E = \{(y, x) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x) \land (-\varepsilon_y \leq y \leq \varepsilon_y) \}$ for $\varepsilon_x, \varepsilon_y > 0$, leaves out points that are small in both components.
An exclusion area along one axis, for example, $E = \{(y, x) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x)\}$ for $\varepsilon_x > 0$, removes points in which one of the components could change sign by a small amount of noise.
A cross-shaped exclusion area, $E = \{(y, x) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x) \lor (-\varepsilon_y \leq y \leq \varepsilon_y) \}$ for $\varepsilon_x, \varepsilon_y > 0$, along both axes accounts for the sign reversal in both components.

@@

\begin{figure}
    \centering
    \begin{subfigure}[t]{.48\textwidth}
    \includegraphics{plots/covid_nowcast/40_cond_prob_lag_7}
    \caption{Conditional ATC plot.}\label{fig:app-covid-cond-prob-7}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{.48\textwidth}
    \includegraphics{plots/covid_nowcast/40_acc_eps_lag_7}
    \caption{ATC ratio over exclusion area size in $\diffx$.}\label{fig:app-covid-atc-ratio-7}
    \end{subfigure}
    \caption[Conditional ATC plot and ATC ratio over exclusion area for the COVID-19 nowcasts.]{Conditional \ac{atc} plot and \ac{atc} ratio over exclusion area for the nowcasts of the seven-day hospitalization rate ILM, RKI, RIVM, and ENS-MED for the horizon seven days in COVID-19 nowcasting.}
    \label{fig:app-covid-cond-prob-atc-ratio-7}
\end{figure}

@@

\section*{Declarations}
\subsection*{Ethics approval and consent to participate}

\subsection*{Consent for publication}

\subsection*{Availability of data and materials}

The datasets analyzed during the study are available in the corresponding databases or repositories (\url{https://covid19nowcasthub.de/}, \url{https://github.com/bahmanrostamitabar/hourly-emergency-care}, and \url{https://physionet.org/content/mimic3wdb/1.0/}).
All code is available in the repository \url{https://github.com/jo-rie/aatc}.

@@

\printbibliography

@@

\begin{figure}
    \centering
    \begin{subfigure}[t]{.48\textwidth}
    \includegraphics{plots/covid_nowcast/40_cond_prob_lag_1}
    \caption{Conditional \ac{atc} plot for horizon one.}\label{fig:app-covid-cond-prob-1}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{.48\textwidth}
    \includegraphics{plots/covid_nowcast/40_cond_prob_lag_14}
    \caption{Conditional \ac{atc} plot for horizon 14.}\label{fig:app-covid-cond-prob-14}
    \end{subfigure}
    \begin{subfigure}[t]{.48\textwidth}
    \includegraphics{plots/covid_nowcast/40_acc_eps_lag_1}
    \caption{\Ac{atc} ratio over exclusion area size in $\diffx$ for horizon one.}\label{fig:app-covid-atc-ratio-1}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{.48\textwidth}
    \includegraphics{plots/covid_nowcast/40_acc_eps_lag_14}
    \caption{\ac{atc} ratio over exclusion area size in $\diffx$ for horizon 14.}\label{fig:app-covid-atc-ratio-14}
    \end{subfigure}
    \caption[Conditional ATC plot and ATC ratio over exclusion area in COVID-19 nowcasting.]{Conditional \ac{atc} plot and \ac{atc} ratio over exclusion area for the COVID-19 nowcast models ILM, RKI, RIVM, and ENS-MED for the horizon seven days. Note the different axis scalings of the ATC ratios for the two horizons. While the ENS-MED model has a rather low \ac{atc} ratio for small exclusion areas, it is increasing fast for larger exclusion areas. The RKI model is among the worst, except for the small negative predicted changes. }
    \label{fig:app-covid-cond-prob-atc-ratio-1-14}
\end{figure}

\begin{figure}
    \centering
    \begin{subfigure}[t]{.48\textwidth}
        \includegraphics{plots/covid_nowcast/60_reliability_diagram_lag_7}
        \caption{Reliability diagram for horizon seven days.} \label{fig:app-covid-reliability-7}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{.48\textwidth}
        \includegraphics{plots/covid_nowcast/60_reliability_diagram_lag_14}
        \caption{Reliability diagram for horizon 14 days.} \label{fig:app-covid-reliability-14}
    \end{subfigure}
    \begin{subfigure}{\textwidth}
        \includegraphics{plots/covid_nowcast/70_prob_hist}
        \caption{Count histogram of the predicted probabilities for the horizon one, seven, and 14 days.} \label{fig:app-covid-prob-hist}
    \end{subfigure}
    \caption[Reliability diagrams for COVID-19 nowcasting models and the horizon seven and 14 days.]{The reliability diagram for the COVID-19 nowcasting models ILM, RIVM, RKI, and ENS-MED for the horizon seven and 14 days.
Additionally, the count of predicted probabilities for the horizons is shown.
    The reliability diagram bins are chosen according to the empirical quantiles of the predicted probabilities.
As the models issue small or large probabilities of increase for the higher horizons, little information on the accuracy of moderate probability predictions is available.}
    \label{fig:app-covid-rel}
\end{figure}
